---
title: "3.Data Visualisation"
author: "HaiXiao Lu"
date: "3/5/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, include = FALSE, message=FALSE, warning=FALSE}
library(tidyverse)
```

# load data 
```{r, echo = FALSE}
mpg
```

## Creating a ggplot
```{r, echo = TRUE}
ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy))
```

Note: The plot shows a negative relationship between engine size(displ) and 
fuel efficiency (hwy). In other words, cars with big engines use more fuel. 

With ggplot2, ggplot() creates a coordinate system that we can add layers to. 
We complete our graph by adding one or more layers to ggplot(). The function
geom_point() adds a layer of points to our plot, which creates a scatterplot. 

### A graphing template
`ggplot(data = <DATA>) + `
    `<GEOM_FUNCTION>(mapping = aes(<MAPPINGS>))`
    
    
#### Exercises

##### Run `ggplot(data = mpg).` what do you see?
```{r, echo = TRUE}
ggplot(data = mpg)
```
    

##### How many rows are in mpg? How many columns?
```{r, echo = TRUE}
dim(mpg)
glimpse(mpg)
```

##### What does the drv variable describe?
```{r}
?mpg
```
drv: the type of drive train, where f = front-wheel drive, 
r = rear wheel drive, 4 = 4wd

##### Make a scatterplot of hwy vs cyl
```{r, echo = TRUE}
ggplot(data = mpg) + 
    geom_point(mapping = aes(x = hwy, y = cyl))
```
    
##### What happens if you make a scatterplot of `class` vs `drv`? why is the plot not useful?
```{r, echo = TRUE}
ggplot(data = mpg) +
    geom_point(mapping = aes(x = class, y = drv))
```

## Aesthetic mappings
You can convey information about your data by mapping the aesthetics in your plot
to the variables in your dataset.
```{r, echo = TRUE}
ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy, color = class))
```

We could have mapped class to the size aesthetic in the same way. 
```{r, echo = TRUE}
ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy, size = class))
```
    
Or we could have mapped class to the `alpha` aesthetic, which controls the transparency of the points, or to the shape aesthetic, which controls the shape of
the points
```{r, echo = TRUE}
# Left
ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy, alpha = class))

# Right
ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy, shape = class))
```
    
The `aes()` function gathers together each of the aesthetic mappings used
by a layer and passes them to the layer's mapping argument. The syntax highlights 
a useful insight about x and y

We can also set the aesthetic properties of our geom manualy. For example, we can make all of the points in our **plot blue**
```{r, echo = TRUE}
ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy), color = 'blue')
```
  
### Exercises

#### What's gone wrong with this code? why are the points not blue?
```{r}
ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy, color = 'blue'))
```

we can change it to:
```{r}
ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy), color = "blue")
```

#### Which variables are continuous? How can you see this information when you 
run mpg?
```{r}
glimpse(mpg)
```

#### Map a continuous variable to color, size and shape, How do these aesthetics behave differently for categorical vs. continuous variables?
```{r, echo = TRUE}
# mapping with year
ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy, color = year))
```

```{r}
# mapping with size
ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy, size = cty))
```

```{r}
ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy, shape =drv))
```


## Facets

One way to add additional variables is with aesthetics. Another way, particularly useful for categorical variables, is to split your plot into **facets**, subplots that each display one subset of the data

The facet your plot by a single variable, use `facet_wrap()`. The first argument of
`facet_wrap()` should be a formular, which you create with `~` followed by a variable name. The variable that you pass to `facet_wrap()` should be discrete
```{r}
ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy)) + 
    facet_wrap(~ class, nrow = 2)
```

To facet your plot on the combination of two varaibles, add `facet_grid()` to your
plot call. The first argument of `facet_grid()` is also a formula. This time the 
formula should contain two variable names separated by a ~
```{r}
ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy)) + 
    facet_grid(drv ~ cyl)
```

If you prefer to not facet in the rows or columns dimension, use a. instead of a variable name, e.g. ` + facet_grid(. ~ cyl)`

### Exercises

#### What happens if you facet on a continuous variable?
```{r}
ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy)) + 
    facet_wrap(~cty, nrow = 3)
```
The continuous variable is converted to a categorical variable, and the plot contains a facet for each distinct value


#### What do the empty cells in plot with `facet_grid(drv ~ cyl) mean? How do they
relate to this plot?
```{r}
ggplot(data = mpg) + 
    geom_point(mapping = aes(x = drv, y = cyl))
```

```{r}
ggplot(data = mpg) + 
    geom_point(mapping = aes(x = hwy, y = cty)) + 
    facet_grid(drv ~ cyl)
```

The empty cells (facets) in this plot are combination of `drv` and `cyl` that have
no observations. These are the same locations in the scatter plot of `drv` and `cyl`
that have no points


#### What plots does the following code make? what does . do?
```{r}
ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy)) + 
    facet_grid(drv ~ .)

ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy)) + 
    facet_grid(. ~ cyl)
```
The `drv ~ .` means facet by values of `drv` on the y-axis
The `.~ drv ` means facet by values of `drv` on the x-axis


#### Take the first faceted plot in this section
```{r}
ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy)) + 
    facet_wrap(~ class, nrow = 2)
```

#### Read `?facet_wrap`. What does `nrow`,`ncol` do? what other options control the layout of the individual panels? Why doesn't facet_grid() have nrow and ncol argument?
```{r}
?facet_wrap
```

nrow, ncol : number of rows and columns

The `nrow` and `ncol` arguments are unnecessaary for `facet_grid()` since the number of unique values of the variables specified in the function determines the number of rows and columns


#### When using facet_grid() you should usually put the variable with 
more unique levels in the columns. why?

There will be more space for columns if the plot is laid out horizontally











